
```{r include=FALSE}
      today <-Sys.Date() 
      
      # today <- as.Date(today, "%d/%m/%Y")
```
---
title: "3-8-2 checks"
author: "Abhishek Singh"
date: "**`r format(today, format="%d-%B-%Y")`**"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(pander)
```

```{r include= FALSE}

date <- Sys.Date()
new_data <- read.csv(paste0(output_folder, "/", date, "_update_3-8-2", ".csv"))

latest_year <- max(new_data$Year)
rm(new_data)

```

## 3-8-2 update summary

Data for `r latest_year` have been created and saved as `r paste0(date, "_3-8-2", ".csv")` 
in the 3-8-2 Output folder.  



### Check configurations
Please check that the information below is as you expect. In particular, pay 
attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
3-8-2/`r input_folder`. 

  
```{r include= FALSE}
filename_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", filename_wb1), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved <- round(difftime(Sys.time(), filename_save_time, units = "days"), 2)


filename_save_time1 <- file.info(paste0(getwd(),"/",input_folder,"/", filename_wb2), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved1 <- round(difftime(Sys.time(), filename_save_time1, units = "days"), 2)


filename_save_time2 <- file.info(paste0(getwd(),"/",input_folder,"/", filename_wb4), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved2 <- round(difftime(Sys.time(), filename_save_time2, units = "days"), 2)




```


**Total household expenditure or income-data  **  

Total household expenditure or income data are in the files **`r filename_wb1`,
`r filename_wb2`,
`r filename_wb4`**.   

This file for **workbook 1** was saved at `r filename_save_time`: It has been 
`r time_since_data_saved` days since the file was saved in 
`r input_folder`.

This file for **workbook 2** was saved at `r filename_save_time1`: It has been 
`r time_since_data_saved1` days since the file was saved in 
`r input_folder`.

This file for **workbook 4** was saved at `r filename_save_time2`: It has been 
`r time_since_data_saved2` days since the file was saved in 
`r input_folder`.

The tabs used are called
**`r household_expd_data`,
`r Age_group_data`,
`r percentage_total_expd_data`,
`r income_decile_group_data`,
`r percent_total_expd_decile_group_data`,
`r household_expd_tenure_data`,
`r household_expd_socioeconomic_data`**. 



### Check data

Because this automation only produces the data for the latest year, it is not 
possible to plot the new data over the old data to check for discrepancies. 
Instead, the person QAing this update should download the source data and spot 
check values the source data against the outputted csv.

### Note
Zero values are greater than 0 but round to 0.
 

### Calculations

No calculations have been performed on this data during this automation.

## Session Info
```{r}
pander(sessionInfo())
```

