---
title: "5-6-1 QA checks"
author: "Ali Campbell"
date: "16/01/2022"
output: html_document

---
*This is only a template and will not produce real output as the variables *
*referred to in the code do not exist - in a real one they will be pulled from*
*the other scripts. For a working QA file similar to this template see 13-2-2*

*Example text is indented, but shouldn't be indented in the real thing*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
library(pander)
```
**Run date and time: `r Sys.time()`**  
  
## Introduction

Short explanation that can include details such as:   
  
- 5-6-1 is a relatively straightforward indicator but contains lots of disaggregations so automating it saves a lot of time in wrangling the data. Since the last manual update we have included further data on region and local authority that benefits from automation.
  
- This file runs some basic checks on the csv output and highlights some specific things that should manually be checked.  

- Make sure that the filenames are all in the correct format

- Check that you have inputted the years variable correctly in config.R

- Is set up so that should all work if you click 'Source' in compile_tables.R

- This automation assumes that the format of Table 7 and Table 17a does not change in future year
  
- Other important info you think the user needs to be aware of.
  
  
## Configurations
in this section list the settings (configurations) and use snippets of code to 
print the user input. The double asterisks make the text bold, while `r` tells 
markdown to read the text inside the back-ticks as code. 
  
  The configurations are:  
  
  - filepath for the input data: **`r paste0(getwd(), "/", input_folder, "/", filename)`** (place .xlsx files here)
  - years that are being read in: **`r years`**
  - filenames being read in: **`r filename_list`**
  - name of the tabs the data are in: **`r tabname_age & tabname_la`** (Note: tabname_la is slightly different from 21/22 onwards compared to earlier years, this will display the tabname for the most recent year read) 
  - row number of the column names; age: **`r header_row_age`**; local authority: **`r header_row_la`** 
  - location of the outputs: **`r output_folder`**  
  
  
## Basic checks
Use this section to highlight anything you think may have gone wrong in the 
code. You may want to use chunks of code within this file to perform some basic 
checks, or you may just pull in variables that were created in the main scripts 
(like with the configurations above). 

- No duplicates in output data: **`r check_output`**


## Plots
The plots are one of the most important parts of the QA file. If anything has 
gone wrong in the automation it will show up on the plots. Below is an example
plots section:  
  
  Please check the plots below for any differences between the live data (which 
  is pulled in from the sdg-data (platform) github repo) and the new data.  
  
  Live data will appear over the top of new data. You should therefore only 
  see points for the live data for previous years and only points for the new data 
  in the years you are updating. 

```{r, echo = FALSE, fig.width=10}

# This example is taken from the 13-2-2 update because it is a straightforward
# one.
#
# Some will be more complex, for example when there are multiple disaggregations.
# See the 3-c-1 QA.Rmd file for an example of more complex QA plots
#
# Note that you may have improved the indicator at the same time as updating it
# when writing the automation. If this is the case, the plots code you write for 
# the first round of checks may need to be slightly different to how it will 
# need to be in future years. So you will need to check the Rmd file works
# with 'live' data that looks like the new data (for testing you could pull the 
# 'live' data from the in progress file once you have put the new csv in there).

live_data <- read.csv('https://raw.githubusercontent.com/ONSdigital/sdg-data/master/data/indicator_5-6-1.csv') %>%
  janitor::clean_names()

new_data <- csv_output %>% # replace csv_formatted with whatever your final csv was called
  mutate(Year = as.integer(Year),
         dataset = "new") %>%
  janitor::clean_names() 
  
# join to the new data
joined_data <- live_data %>% 
  mutate(dataset = "live") %>% 
  bind_rows(new_data)

# plot the data
plot <- joined_data %>% 
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point() +
  geom_line(aes(linetype = dataset)) +
  facet_wrap(vars(gas),
             nrow = 2) +
  theme_bw() 

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

Finally, it is sometimes hard to see where new numbers don't match live numbers
especially if there are lots of disaggregations. In such cases a table of 
differences is also useful. The following example is taken from 13-2-2:
  
  At the time of writing this code (June 2022), the following note is given in the
  source notes tab: 'in these statistics the entire time series going back to 
  1990 is revised each year to take account of methodological improvements'.  
  
  **If there are differences (see table below) please check that this still stands**

```{r, , echo = FALSE}
comparison <- joined_data %>% 
  # select the columns that are in both the live and new data
  # Columns that have the same entry in every row (e.g. units), or
  # columns like geocode and obs status don't need to go in here.
  select(year, gas, value, dataset) %>% 
  pivot_wider(names_from = dataset,
              values_from = value) %>% 
  mutate(difference = new - live) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```
  
  The table below shows all instances where the old value does not match the new 
  value. It is sorted so the largest absolute differences appear first, but can
  be sorted in other ways using the arrows at the top.
```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

