---
title: "5-6-1 QA checks"
author: "Ali Campbell"
date: "16/01/2022"
output: html_document

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
library(pander)
```
**Run date and time: `r Sys.time()`**  
  
## Introduction
  
- This file runs some basic checks on the csv output and highlights some specific things that should manually be checked.  

- 5-6-1 is a relatively straightforward indicator but contains lots of disaggregations so automating it saves a lot of time in wrangling the data. Since the last manual update we have included further data on region and local authority that benefits from automation.

- Make sure that the filenames are all in the correct format

- Check that you have inputted the years variable correctly in config.R

- Is set up so that should all work if you click 'Source' in compile_tables.R (make sure working directory is set to 5-6-1 in sdg_data_updates)

- This automation assumes that the format of Table 7 and Table 17a does not change in future years
  
  
## Configurations
  The configurations are:

  - filepath for the input data: **`r paste0(getwd(), "/", input_folder)`** (place .xlsx files here)
  - years that are being read in: **`r years`**
  - filenames being read in: **`r filename_list`**
  - name of the tabs the data are in: **`r tabname_age`** and **`r tabname_la`** (Note: the tabname for the local authority data is 'Table 17' for years earlier than 21-22, this will display the tabname for the most recent year read) 
  - row number of the column names; age: **`r header_row_age`**; local authority: **`r header_row_la`** 
  - location of the outputs: **`r output_folder`**  
  
  
## Basic checks
Both should be true:

  - No duplicates in output rows: **`r check_output`**
  - No duplicates (not including values): **`r check_wo_values`**

## Plots
Please check the plots below for any differences between the live data (which 
is pulled in from the sdg-data (platform) github repo) and the new data.  
Live data will appear over the top of new data. You should therefore only 
see points for the live data for previous years and only points for the new data 
in the years you are updating.

```{r, echo = FALSE, fig.width=10}

# This example is taken from the 13-2-2 update because it is a straightforward
# one.
#
# Some will be more complex, for example when there are multiple disaggregations.
# See the 3-c-1 QA.Rmd file for an example of more complex QA plots
#
# Note that you may have improved the indicator at the same time as updating it
# when writing the automation. If this is the case, the plots code you write for 
# the first round of checks may need to be slightly different to how it will 
# need to be in future years. So you will need to check the Rmd file works
# with 'live' data that looks like the new data (for testing you could pull the 
# 'live' data from the in progress file once you have put the new csv in there).

# Read in the data currently on the website
live_data <- read.csv('https://raw.githubusercontent.com/ONSdigital/sdg-data/master/data/indicator_5-6-1.csv') %>%
  janitor::clean_names()

#Change names to match
live_data$types_of_contraception <- gsub("Other methods", "Other methods [age]", as.character(live_data$types_of_contraception))

# Read in the new data 
new_data <- csv_output %>% 
  mutate(mutate(across(where(is.double), as.character)),
         dataset = "new") %>%
  # These two lines reformat the years back into the live format
  mutate(Year = str_replace_all(Year, "-", "/")) %>%
  mutate(Year = paste("20", Year, sep = "")) %>%
  janitor::clean_names() 
  
# join to the new data
joined_data <- live_data %>% 
  mutate(across(where(is.double), as.character), dataset = "live") %>%
  rename(type_of_contraception = types_of_contraception) %>%  #changes name to match
  bind_rows(new_data)

# Remove NAs local authorities/regions (la_data not included prior to 2019/20)
joined_data <- joined_data %>%
  mutate(region_name = ifelse(is.na(region_name), "", region_name)) %>%
  mutate(la_name = ifelse(is.na(la_name), "", la_name))

# Changes names to match
joined_data$type_of_contraception <- gsub("Natural family planning incl. rhythm method", "Natural family planning", as.character(joined_data$type_of_contraception))

# Convert values to numeric so can be subtracted from each other later
joined_data$value <- as.numeric(joined_data$value)

# PLOTS
# There are too many possible time series to plot here but below are a few examples you can amend if there seems to be a discrepancy shown in the table above. 
# Age
plot_age <- joined_data %>%
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point() +
  geom_line(aes(linetype = dataset)) +
  facet_wrap(~ age,
             nrow = 4) +
  theme_bw() +
  ggtitle("Age")

# Main method of contraception
plot_method <- joined_data %>%
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point() +
  geom_line(aes(linetype = dataset)) +
  facet_wrap(~ main_method_of_contraception, nrow = 3) +
  theme_bw() +
  ggtitle("Main method of contraception")

# Type of contraception
plot_type <- joined_data %>%
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point() +
  geom_line(aes(linetype = dataset)) +
  facet_wrap(~ type_of_contraception, nrow= 6) +
  theme_bw() +
  ggtitle("Type of contraception")

# Region name
plot_region <- joined_data %>%
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point() +
  geom_line(aes(linetype = dataset)) +
  facet_wrap(~ region_name, nrow = 5) +
  theme_bw() +
  ggtitle("Region")
```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot_age))
```

```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot_method))
```

```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot_type))
```

```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot_region))
```



```{r, , echo = FALSE}
# joined_data$value <- as.numeric(joined_data$value)
joined_data <-  joined_data %>%
    select(all_of(c("year", "region_name", "la_name", "age", "main_method_of_contraception", "type_of_contraception",  "unit_measure", "value", "dataset", "observation_status", "unit_multiplier")))
  

comparison <- joined_data %>% 
  # select the columns that are in both the live and new data
  # Columns that have the same entry in every row (e.g. units), or
  # columns like geocode and obs status don't need to go in here.
  select(year, age, region_name, la_name, main_method_of_contraception, type_of_contraception, value, dataset) %>% 
  pivot_wider(names_from = dataset,
              values_from = value) %>%
  mutate(difference = new - live) %>%
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))
```
  
  The table below shows all instances where the old value does not match the new 
  value. It is sorted so the largest absolute differences appear first, but can
  be sorted in other ways using the arrows at the top.
```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

