---
title: "template QA"
author: "Emma Wood"
date: "29/06/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
```
**Run date and time: `r Sys.time()`**  
  
## Introduction

- 1-a-2 is a very simple automation which calculates a single percentage for each year

- This file runs some basic checks on the csv output and highlights some specific 
things that should manually be checked. 
  
- Before putting the csv in the indicator file please check in the source data if the most recent 
figures are provisional. If they are, in the indicator csv file change the 
observation status to 'Provisional'.
  
- Please check the configs below and change them if not correct
  
## Configurations
in this section list the settings (configurations) and use snippets of code to 
print the user input. The double asterisks make the text bold, while `r` tells 
markdown to read the text inside the back-ticks as code. 

  The configurations are:  
  
  - filepath for the input data: **`r paste0(getwd(), "/", input_folder, "/", filename)`**
  - name of the tab the data are in: **`r tab_name`**  
  - row number of the column names: **`r header_row`** 
  - location of the inputs **`r input_folder`**
  - location of the outputs: **`r output_folder`**
  - most recent year being read in: **`r most_recent_year`**
  
  
## Basic checks
- No duplicates present: **`r check_all`**

## Plots
Please check the plots below for any differences between the live data (which 
is pulled in from the sdg-data (platform) github repo) and the new data.  
  
Live data will appear over the top of new data. You should therefore only 
see points for the live data for previous years and only points for the new data 
in the years you are updating. 

```{r, echo = FALSE, fig.width=10}
live_data <- read.csv('https://raw.githubusercontent.com/ONSdigital/sdg-data/master/data/indicator_1-a-2.csv') %>%
  janitor::clean_names()

new_data <- csv_output %>% 
  mutate(,
         dataset = "new") %>%
  janitor::clean_names() 
  
# join to the new data
joined_data <- live_data %>% 
  mutate(dataset = "live") %>% 
  bind_rows(new_data)

# plot the data
plot <- joined_data %>% 
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point(aes(shape = dataset)) +
  geom_line(aes(linetype = dataset)) +
  scale_shape_manual(values = c(15, 19)) +
  theme_bw() 

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

**If there are differences (see table below) please check that this still stands**

```{r, , echo = FALSE}
comparison <- joined_data %>% 
  # select the columns that are in both the live and new data
  # Columns that have the same entry in every row (e.g. units), or
  # columns like geocode and obs status don't need to go in here.
  select(year, value, dataset) %>% 
  pivot_wider(names_from = dataset,
              values_from = value) %>% 
  mutate(difference = new - live) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```
  
  The table below shows all instances where the old value does not match the new 
  value. It is sorted so the largest absolute differences appear first, but can
  be sorted in other ways using the arrows at the top.
```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

