---
title: "15-7-1 and 15-c-1 checks"
author: "Katie Uzzell"
date: "22/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
```

```{r include= FALSE}
source('config.R')

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/15-7-1.csv")

date <- Sys.Date()
new_data <- read.csv(paste0("Output/", date, "_15-7-1_and_15-c-1", ".csv")) 

latest_year <- max(new_data$Year)

```

## 15-7-1 and 15-c-1 update summary

Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_15-7-1_and_15-c-1", ".csv")` in the 15-1-1 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay 
attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
15-7-1_and_15-c-1/`r input_folder`.  
  
```{r include= FALSE}
filename_save_time <- file.info(paste0(getwd(),"/Input/", filename), full.names = TRUE)$ctime[1]
time_since_data_saved <- round(difftime(Sys.time(), filename_save_time, units = "days"), 2)

```
<br>

**Woodland data from Forest Research**  
Woodland area data are in the file `r filename`.   

This file was saved at `r filename_save_time`: It has been 
`r time_since_data_saved` days since the file was saved in 
`r input_folder`  

The tab used for 'all woodland' areas is called
`r tabname`. 
    
<br>


### Check data
The charts below show the new data (just created) plotted on top of the old data,
which has been downloaded from the live site. If there are unexpected 
discrepancies in the lines please investigate the reasons for these discrepancies.
    
```{r, echo=FALSE, fig.width=7, fig.height=5}
new_data_to_plot <- mutate(new_data, dataset = "2 new")
old_data_to_plot <- mutate(old_data, dataset = "1 old") 

status_list <- unique(new_data_to_plot$Sustainably.managed.status)

new_data_to_plot %>%
    bind_rows(old_data_to_plot) %>% 
    ggplot(data = .,
           aes(x = Year,
               y = Value,
               colour = dataset)) +
    geom_point() +
    geom_line() +
    facet_grid(Country ~ Sustainably.managed.status) +
    theme_bw() 

```

### Calculations

Q1, Q2, Q3, and Q4 were summed to give a total for each year. 

The calculations used to create these figures was:
```
seizures_main_data_totals <- aggregate(x = seizures_main_data_totals[ , colnames(seizures_main_data_totals) != "Year"],
            by = list(seizures_main_data_totals$Year),
            FUN = sum) 
```
Warning: This is copied from `update_15-7-1_and15-c-1.R` and will not dynamically update 
if that code is changed.

