---
title: "3-2-2 QA"
author: "Emma Wood"
date: "27/09/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
```
**Run date and time: `r Sys.time()`**  
   
## Introduction

3-2-2 is a relatively straightforward indicator but there are a lot of potential
disaggregations so it has been automated to speed up the process.    
    
This file runs some basic checks and gives information about the update.

    
## Configurations

Configurations for this run are as follows:       
    
- Input filepath: **`r paste0(getwd(), "/", input_folder, "/", filename)`**
- Tab name for country of occurrence by sex data: **`r country_of_occurrence_by_sex_tab_name`**    
- Tab name for area of residence data: **`r area_of_residence_tab_name`**    
- Tab name for birthweight by mother age data: **`r birthweight_by_mum_age_tab_name`**  
- Tab name for mother's country of birth: **`r country_of_birth_tab_name`**
- Tab name for ethnicity: **`r ethnicity_tab_name`**
- The first row of headings for each of these data sets are set as rows 
`r first_header_row_country_by_sex`, `r first_header_row_area_of_residence`, 
`r first_header_row_birthweight_by_mum_age`, and 
`r first_header_row_country_of_birth` respectively.    
`r first_header_row_ethnicity` respectively.    
- Data have been rounded to **`r decimal_places` decimal places**.   

## Potential errors

### Year

Year is taken from the information above the headings in the source spreadsheet.
If the cell containing the year contains a superscript straight after the year, 
year may not be picked up.    

```{r, include = FALSE}
country_sex_year <- unique(clean_csv_data_country_by_sex$Year) 
area_of_residence_year <- unique(clean_csv_data_area_of_residence$Year) 
birthweight_age_year <- unique(clean_csv_data_birtweight_by_mum_age$Year) 
if(include_country_of_birth == TRUE) {
  country_of_birth_year <- unique(clean_csv_data_country_of_birth$Year)
}


blank_years <- c("Years were not identified in the following tabs in the source spreadsheet:")
years <- c()
if(country_sex_year == "") {
  blank_years <- c(blank_years, country_of_occurrence_by_sex_tab_name)
} else {
    years <- c(years, country_sex_year)
}

if(area_of_residence_year == "") {
  blank_years <- c(blank_years, area_of_residence_tab_name)
} else {
    years <- c(years, area_of_residence_year)
}

if(birthweight_age_year == "") {
  blank_years <- c(blank_years, birthweight_by_mum_age_tab_name)
} else {
    years <- c(years, birthweight_age_year)
}
if(include_country_of_birth == TRUE){
  if(country_of_birth_year == "") {
    blank_years <- c(blank_years, country_of_birth_tab_name)
  } else {
    years <- c(years, country_of_birth_year)
  }
}


if(length(blank_years) == 1){
  blank_years <- "Year was identified in all tabs"
} else {
  blank_years <- paste(c(blank_years, "Please check this/these tabs to see why, **manually add the correct year to the blank cells**, and check if any notes need to be added to the metadata.  Please also check that the year given is as expected"), collapse = '<br />')

}

if(length(unique(years)) == 1) {
  potential_year_issue <- ""
} else {
  potential_year_issue <- "Different years were found in different tables, but 
  they should all be the same. Please investigate."
}
```
    
`r blank_years`    
    
`r potential_year_issue`    
    
### Blank values
Blank values due to suppression have been retained in the output csv as this seems to be in line with
SDMX options. This is most often the case in cross-disaggregtaions for more extreme groups.
For example very low birthweights, and the oldest and youngest age groups.  
  
This table shows rows for which we would definitely expect to have a value but don't 
(where it is not a cross-disaggregation, birthweight is not "unlinked deaths" and 
neither age nor birthweight are "not stated"). If this table contains unexpected disaggregations, 
check whether the value is present in the original data.
  
```{r, echo = FALSE}
no_value_rows_filtered <- no_value_rows %>% 
  filter(Birthweight != "Unlinked deaths" &
           Birthweight != "Not stated" & 
           Age != "Not stated" &
           `Country of birth` != "Other/Not stated" &
           `Country of birth` != "Born in Antarctica or Oceania") %>% 
  mutate(unexpected = case_when(
    Sex != "" & Birthweight == "" & Age == "" & `Neonatal period` == "" ~ TRUE,
    Birthweight != "" & Sex == "" & Age == "" & `Neonatal period` == "" ~ TRUE,
    Age != "" & Birthweight == "" & Sex == "" & `Neonatal period` == "" ~ TRUE,
    `Neonatal period` != "" & Birthweight == "" & Age == "" & Sex == "" ~ TRUE,
    TRUE ~ FALSE)) %>% 
  filter(unexpected == TRUE) %>% 
  select(-unexpected)

datatable(no_value_rows_filtered, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```
The table below shows **all** rows for which there is no value given (these rows do not appear in the csv). 
```{r, echo = FALSE}
datatable(no_value_rows, 
          rownames = FALSE, 
          filter = "top", 
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))

```
### Calculations
For birthweight and mother age, and for country of mother's birth, 
number of late neonatal deaths are calculated as
'number of neonatal deaths - number of early neonatal deaths.   

The rates for early and late neonatal deaths per 1,000 live births are then
calculated as 'number of early [or late] neonatal deaths / (number of live births/1,000)',
where the numerator is greater than or equal to 3.

As a preliminary check, the same calculation is run for all neonatal deaths and
this is compared to the published rate for all neonatal deaths. If these two match,
we can be confident the functions used to calculate rate are working as they should.
If all or most do not match, this indicates a problem that needs to be investigated.  
  
N.B. there may be occasional mismatches due to differences in rounding. The `round()` 
function rounds .5 to the even number, as always rounding up (as Excel does) leads to accumulation of error when errors are averaged).   
  
See tables below for mismatches:
```{r, echo = FALSE}

if(rate_mismatches_weight_age > 0){
  
table1 <- calculations_weight_age %>% 
  filter(Rates_Neonatal_check != Neonatal_rate) %>% 
    distinct(birthweight, mother_age, Rates_Neonatal_check, Neonatal_rate)

datatable(table1, 
          rownames = FALSE, 
          filter = "top", 
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))

} else {"there are no mismatches for birthweight and age of mother"}

if(include_country_of_birth == TRUE){

    if(rate_mismatches_country_of_birth > 0){
    
    table2 <- calculations_country_of_birth %>% 
      filter(Rates_Neonatal_check != Neonatal_rate) %>% 
      distinct(mother_country, Rates_Neonatal_check, 
               Neonatal_rate)
    
    datatable(table2, 
              rownames = FALSE, 
              filter = "top", 
              options = list(dom = 'tp', 
                             pageLength = 5,
                             columnDefs = list(list(
                               className = 'dt-center'))))
    
  } else {print("there are no mismatches for mother's country of birth")}
  
}

```

Further to this, the rates of early and late neonatal deaths should sum to the 
rate of all neonatal death (which is in the publication). The table below shows 
rows where the sum of the early and late neonatal rates is different to the 
published neonatal rate by 0.1 or more, ordered by largest difference:

```{r, echo = FALSE}
wide_data <- csv_data %>% 
  mutate(`Neonatal period` = ifelse(`Neonatal period` == "", 
                                    "Neonatal", `Neonatal period`)) %>% 
  pivot_wider(names_from = `Neonatal period`,
              values_from = Value) %>% 
  mutate(sum_early_late = as.numeric(`Early neonatal`) + 
           as.numeric(`Late neonatal`)) %>% 
  mutate(difference = sum_early_late - as.numeric(Neonatal)) %>% 
  filter(difference >= 0.1) %>% 
  arrange(desc(difference)) 

datatable(wide_data, 
          rownames = FALSE, 
          filter = "top", 
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

For most disaggregations late neonatal rate (where the count of 
late neonatal deaths is 3 or greater) is calculated as:  
(number of neonatal deaths - number of early neonatal deaths) /  
(number of live births / 1000)  
However, where country of occurrence is Northern Ireland, 
the rate for late neonatal deaths is calculated as 
'neonatal death rate - early neonatal death rate' (where the count of 
late neonatal deaths is 3 or greater). the number of live births *doesn't* include non-residents,
but the number of deaths *does* include non-residents. The published rates use number of 
live births including non-residents so that the numerator and denominator are comparable.  
The reason I haven't used this simpler calculation for all disaggregations is twofold:  
1. The former calculation is more precise because of rounding in the rates  
2. Early neonatal death rate is not available for birthweight by age of mother    


## Plots
  
Plots of the data are shown below. If you see anything unexpected in the plots,
please manually check the data are correct, and edit the code if necessary.

```{r, echo = FALSE, fig.width=10}
old_data <- read.csv("Y:/Data Collection and Reporting/Jemalex/CSV/indicator_3-2-2.csv", 
                     encoding = "UTF-8") %>% 
  mutate(across(where(is.factor), as.character)) %>% 
  mutate(Region = case_when(
    Region == "East of England" ~ "East", 
    Region == "Yorkshire and the Humber" ~ "Yorkshire and The Humber",
    TRUE ~ as.character(Region)),
    Value = round(Value, decimal_places),
    dataset = "current") %>% 
  rename(Birthweight = Birthweight..grams.,
         `Country of birth` = Country.of.birth,
         `Unit multiplier` = Unit.multiplier,
         `Neonatal period` = Neonatal.period,
         `Observation status` = Observation.status) 

csv_data <- csv_data %>% 
  # mutate(Year = year) %>% 
  mutate(Year = as.integer(Year),
         dataset = "new")

all_data <- old_data %>% 
  bind_rows(csv_data) %>% 
  mutate(`Neonatal period` = ifelse(is.na(`Neonatal period`), "", `Neonatal period`)) %>% 
  mutate(`period and dataset` = paste(dataset, `Neonatal period`))

# country by sex
country_sex_plot <- all_data %>% 
  filter(Birthweight == "" &
         Region == "" &
           Age == "" &
           (`Country of birth` == "" | is.na(`Country of birth`)) &
           (`Ethnic group` == "" | is.na(`Ethnic group`))) %>% 
  ggplot(data = .,
       aes(x = Year,
           y = Value)) +
  geom_point(aes(colour = `period and dataset`)) +
  geom_line(aes(colour = `period and dataset`)) +
  facet_grid(Sex ~ Country) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  ggtitle("country of occurrence by sex")


age_plot <- all_data %>% 
  filter(Region == "" &
           Sex == "" &
           Birthweight == "" &
           Age != "Not stated" &
           (Country == "England and Wales" & Age != "") &
           (`Ethnic group` == "" | is.na(`Ethnic group`))) %>% 
  ggplot(data = .,
       aes(x = Year,
           y = Value)) +
  geom_point(aes(colour = `period and dataset`)) +
  geom_line(aes(colour = `period and dataset`)) +
  facet_grid(. ~ Age) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  ggtitle("Age of mother")

birthweight_plot <- all_data %>% 
  filter(Region == "" &
           Sex == "" &
            Age == "" & 
           (`Country of birth` == "" | is.na(`Country of birth`)) &
           (`Ethnic group` == "" | is.na(`Ethnic group`))) %>% 
  ggplot(data = .,
       aes(x = Year,
           y = Value)) +
  geom_point(aes(colour = `period and dataset`)) +
  geom_line(aes(colour = `period and dataset`)) +
  facet_grid(. ~ Birthweight) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  ggtitle("Birthweight")

region_plot <- all_data %>% 
  filter(Birthweight == "" &
           Sex == "" &
           Age == "" &
           Country == "England" & Region != "" &
           (`Country of birth` == "" | is.na(`Country of birth`)) &
           (`Ethnic group` == "" | is.na(`Ethnic group`))) %>% 
  ggplot(data = .,
       aes(x = Year,
           y = Value)) +
  geom_point(aes(colour = `period and dataset`)) +
  geom_line(aes(colour = `period and dataset`)) +
  facet_grid(. ~ Region) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
  ggtitle("Regions of England")

if(include_country_of_birth == TRUE) {
  country_of_birth_plot <- all_data %>% 
    filter(`Country of birth` != "" &
             !is.na(`Country of birth`) &
           (`Ethnic group` == "" | is.na(`Ethnic group`))) %>% 
    ggplot(data = .,
           aes(x = Year,
               y = Value)) +
    geom_point(aes(colour = `period and dataset`)) +
    geom_line(aes(colour = `period and dataset`)) +
    facet_grid(. ~ `Country of birth`) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
    ggtitle("Country of mother's birth")
}

if (pre_2020_data == FALSE) { # beacuse there is no script for ethnicity in the older data
  ethnicity_plot <- all_data %>% 
    filter(
      `Ethnic group` != "" & !is.na(`Ethnic group`) &
        Birthweight == "" &
        Sex == "" &
        Age == "" &
        Country %in% c("England", "England and Wales") & 
        Region == "" &
        (`Country of birth` == "" | is.na(`Country of birth`))) %>% 
    ggplot(data = .,
           aes(x = Year,
               y = Value)) +
    geom_point(aes(colour = `period and dataset`)) +
    geom_line(aes(colour = `period and dataset`)) +
    facet_grid(Country ~ `Ethnic group`) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1)) +
    ggtitle("Ethnic group")
}

suppressMessages(suppressWarnings(print(country_sex_plot)))
suppressMessages(suppressWarnings(print(region_plot)))
suppressMessages(suppressWarnings(print(age_plot)))
suppressMessages(suppressWarnings(print(birthweight_plot)))
if(include_country_of_birth == TRUE){
  suppressMessages(suppressWarnings(print(country_of_birth_plot)))
}
if(pre_2020_data == FALSE) {
    suppressMessages(suppressWarnings(print(ethnicity_plot)))
}

```






