title: "6-2-1 Checks"
author: "Michael Nairn and Tom McNulty
date: "04/07/2023"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(DT)
library(pander)
```



   
## 6-2-1 update summary

This file runs some basic checks and gives information about the update.

Before putting the csv in the indicator file please check if the most recent figures are provisional. Check this in the source data. If they are, in the indicator csv file change the observation status to 'Provisional'.



```{r include= FALSE}
old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/6-2-1.csv") %>% 
  rename(`Local Authority` = `Local.Authority`) %>%
  rename(`Unit measure` = "Units") %>%
  rename(`Observation status` = Observation.status) %>%
  mutate(dataset = "old") %>%
  select(Year, Series, Region, `Local Authority`, Sex,
         `Unit measure`, `Observation status`, Value, dataset)

# reformat the series 
old_data$Series <- gsub("[(%)]", "" , old_data$Series)
old_data$Series <- str_trim(old_data$Series, "right")



date <- Sys.Date()
new_data <- read.csv(paste0("Output/", date, "_6-2-1", ".csv")) %>% 
  rename(`Local Authority` = `Local.Authority`) %>%
  rename(`Unit measure` = "Unit.measure") %>%
  rename(`Observation status` = Observation.status) %>%
  mutate(dataset = "new") %>%
    select(Year, Series, Region, `Local Authority`, Sex,
         `Unit measure`, `Observation status`, Value, dataset)

 

# join to the new data
joined_data <- old_data %>% 
  bind_rows(new_data) 



```





```{r, , echo = FALSE}
comparison <- joined_data %>% 
  na.omit(Value) %>% # replace the NULL returns with NA
  pivot_wider(names_from = dataset,
              values_from = Value)


comparison <- comparison %>% 
  mutate(across("old":"new", ~replace(., lengths(.) == 0, ""))) %>%
  mutate(new = as.character(new)) %>%
  mutate(new = as.numeric(new)) %>%
  mutate(old = as.character(old)) %>%
  mutate(old = as.numeric(old))

comparison <- comparison %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```

## Table of discrepancies
The table below shows all instances where the old value does not match the new 
value. It is sorted so the largest absolute differences appear first, but can
be sorted in other ways using the arrows at the top.

If there are differences, visible in the table and plots below, please investigate these.
It is possible the data source has update its back time series, so the difference
could be genuine. This is particularly the case for June/July 2023, when the source has 
changed. 



```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```
  



## Headline Plots
Please check the plots below for any differences between the live and new data.
Live data will appear over the top of new data, so if you don't see points for
the new data on dates where there is also live data this is a good thing. 

For the June 2023 update there are a number of new disaggregations not previously 
available on the live site. Therefore, there will be a lot of incomparable 
data points on the charts. 



```{r, echo = FALSE, fig.width=10}

# Headline
headline_plot <- joined_data %>% 
  filter(Series == "All publicly funded schools" & Region == "" & `Local Authority` == "" 
        & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Headline")


# All school types by region
region_plot <- joined_data %>% 
  filter(`Local Authority` == ""  & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Series ~ Region) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("All school types by region")

# All school types by sex
sex_plot <- joined_data %>% 
  filter(`Local Authority` == ""  & Region == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Series ~ Sex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("All school types by sex")

```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(headline_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(region_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(sex_plot))
```


## Disaggregation Plots

There are too many possible time series to plot here 
but below are a few examples you can amend if there seems to be a discrepancy. In 
particular for any discrepancies shown in the table. 
You may need to change the filters in the plots.

For example, selecting specific local authorities by editing the "filter" command to 
include the local authorities you wish to compare. 
  E.g.   filter(`Local authority` == "Barking and Dagenham" | 
                `Local authority` == "Barnsley")



```{r, echo = FALSE, fig.width=10}

# Primary by Sex and Region
primary_by_sex_and_region_plot <- joined_data %>% 
  filter(Series == "Primary"& `Local Authority` == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Region ~ Sex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Primary by Sex and Region")


```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(primary_by_sex_and_region_plot))
```



## Session Info
```{r}
pander(sessionInfo())
```
