---
title: "8-10-1 checks"
author: "Katie Uzzell"
date: "26/07/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(pander)
```

```{r include= FALSE}

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/8-10-1.csv")

date <- Sys.Date()
new_data <- read.csv(paste0(output_folder, "/", date, "_update_8-10-1", ".csv"))

latest_year <- max(new_data$Year)

```

## 8-10-1 update summary

Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_8-10-1", ".csv")` in the 8-10-1 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
8-10-1/`r input_folder`.  

### Check data
The charts below show the new data (just created) plotted on top of the old data, which has been downloaded from the live site. It is possible there will have been revisions to the backseries, so small discrepancies may be expected. If there are unexpected or large discrepancies between the old and new data, please investigate the reasons for these discrepancies.
    
```{r, echo=FALSE, fig.width=7, fig.height=15}
new_data <- new_data %>%
mutate(Year = as.integer(Year),
         dataset = "new")

new_data$Value <- as.numeric(new_data$Value)

joined_data <- old_data %>%
  mutate(dataset = "old") %>%
  bind_rows(new_data)
 
plot <- joined_data %>%
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
    #scale_x_continuous(breaks = seq(min(year(joined_data$Year)), max(year(joined_data$Year)), by = 3))+
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7))

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

### Calculations

Rate was calculated using the following formula:

((Number of banks + Number of building societies) / Population)) * 100000

## Session Info
```{r}
pander(sessionInfo())
```

