---
title: "8-10-1 checks"
author: "Katie Uzzell"
date: "26/07/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(pander)
```

```{r include= FALSE}

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/8-10-1.csv")

date <- Sys.Date()

new_data <- read.csv(paste0(output_folder, "/", date, "_update_8-10-1", ".csv"))

latest_year <- max(new_data$Year)

```

## 8-10-1 update summary

Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_8-10-1", ".csv")` in the 8-10-1 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
8-10-1/`r input_folder`.  

## **Check data**

Please check the table below for any differences between the live and new data. If there are differences visible in the table below, please investigate these.

```{r include=FALSE}

comparison <- joined_data %>% 
  select(Year, Country, Region, Local.Authority, Value, dataset) %>%
  pivot_wider(names_from = dataset,
              values_from = Value)

comparison <- comparison %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

### Calculations

Rate was calculated using the following formula:

((Number of banks + Number of building societies) / Population)) * 100000

## Session Info
```{r}
pander(sessionInfo())
```

