
```{r include=FALSE}
      today <-Sys.Date() 
      
      # today <- as.Date(today, "%d/%m/%Y")
```
---
title: "8-6-1 checks"
author: "Abhishek Singh"
date: "`r format(today, format="%d-%B- %Y") `"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
```

```{r include= FALSE}

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/8-6-1.csv")


# old_data$Series = str_replace(old_data$Series, "Proportion of youth not in education, employment or training", "Yearly")

old_data$Series = gsub( "Proportion of youth not in education, employment or training (quarterly)",
                        "Quaterly", old_data$Series, fixed=TRUE)

old_data$Series = gsub("Proportion of youth not in education, employment or training",
                       "Yearly", old_data$Series, fixed=TRUE)

old_data$Age = gsub(".", "People", old_data$Age, fixed=TRUE)
old_data["Sex"][is.na(old_data["Sex"])] <- "People"


old_data$`Seasonally.adjusted.or.not` = str_replace(old_data$`Seasonally.adjusted.or.not`, "Seasonally adjusted", "SA")
old_data$`Seasonally.adjusted.or.not` = str_replace(old_data$`Seasonally.adjusted.or.not`, "Not seasonally adjusted", "NSA")






date <- Sys.Date()
new_data <- read.csv(paste0(output_folder, "/", date, "_update_8-6-1", ".csv"))

# new_data$Series = str_replace(new_data$Series, "Proportion of youth not in education, employment or training (quarterly)", "Quarterly")
# new_data$Series = str_replace(new_data$Series, "Proportion of youth not in education, employment or training", "Yearly")

new_data$Series = gsub("Proportion of youth not in education, employment or training (quarterly)",
                       "Quaterly", new_data$Series, fixed=TRUE)

new_data$Series = gsub("Proportion of youth not in education, employment or training",
                       "Yearly", new_data$Series, fixed=TRUE)
# new_data$Age = gsub(".", "People", new_data$Age, fixed=TRUE)
new_data["Sex"][is.na(new_data["Sex"])] <- "People"

new_data$`Seasonally.adjusted.or.not` = str_replace(new_data$`Seasonally.adjusted.or.not`, "Seasonally adjusted", "SA")

new_data$`Seasonally.adjusted.or.not` = str_replace(new_data$`Seasonally.adjusted.or.not`, "Not seasonally adjusted", "NSA")



latest_year <- max(new_data$Year)

```

## 14-2-1 update summary

Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_8-6-1", ".csv")` in the 8-6-1 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay 
attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
8-6-1/`r input_folder`.  
  
```{r include= FALSE}
filename_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", filename), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved <- round(difftime(Sys.time(), filename_save_time, units = "days"), 2)

```

**Proportion of youth (aged 15–24 years) not in education, employment or training**  

Protected areas data are in the file `r filename`.   

This file was saved at `r filename_save_time`: It has been 
`r time_since_data_saved` days since the file was saved in 
`r input_folder`  

The tab used is called
`r tabname`. 






### Check data




The charts below show the new data (just created) plotted on top of the old data,
which has been downloaded from the live site. If there are unexpected 
discrepancies in the lines please investigate the reasons for these discrepancies.
    
```{r, echo=FALSE, fig.width=70, fig.height=15}
new_data <- new_data %>%
mutate(Year = as.character(Year),
         dataset = "new")

# For review, how to plot Q1 til Q4 ??



new_data$Value <- as.numeric(new_data$Value)

joined_data <- old_data %>%
  mutate(dataset = "old") %>%
  bind_rows(new_data)

Yearly_plot_data  <- joined_data %>% 
  filter(Series == "Yearly") 



# get yearly data from joined
Yearly_data <- joined_data %>%
  filter(Series == "Yearly")

Yearly_data <- Yearly_data %>% 
      mutate(Year = as.numeric(Year), Value = as.numeric(Value))







# Yearly plot
Yearly_plot <- Yearly_data %>%
  # filter(Series == "Yearly") %>%
  # joined_data$Year = as.integer(joined_data$Year) %>%
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (Yearly_data$Year):max (Yearly_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(~Series ~Seasonally.adjusted.or.not ~ Age ~ Sex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Proportion of youth not in education, employment or training by Year")
  

# Quaterly Plot
 
plot <- joined_data %>%
  filter(Series == "Quaterly") %>%
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  # scale_y_continuous(limits=c(min(joined_data$Value), max(joined_data$Value))) +
  geom_point(aes(shape = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(~Series ~Seasonally.adjusted.or.not ~ Age ~ Sex,scales = "free_x", space='free_x') +
  scale_x_discrete(expand = c(0, 0.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 3.5)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Proportion of youth not in education, employment or training Quaterly")

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(Yearly_plot))
```

















Warning: This is copied from `update_8-6-1.R` and will not dynamically update 
if that code is changed.
