---
title: "3-2-1 checks"
author: "Ali Campbell"
date: "27/02/2023"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(ggplot2)
library(DT)
library(janitor)
library(pander)
```
**Run date and time: `r Sys.time()`**  
  
## Introduction

Short explanation that can include details such as:   
  
- 3-2-1 data comes in two large tables that need to be formatted down and have a death
rate per 1000 live births calculated
  
- This file runs some basic checks on the csv output and highlights some specific 
things that should manually be checked.
  
## Configurations
  The configurations are:  
  
  - filepath for the input data: **`r paste0(getwd(), "/", input_folder, "/", filename)`**
  - name of the tab the data are in; England & Wales: **`r tab_name_EW`**, UK: **`r tab_name_UK`**
  - row number of the column names: E&W: **`r header_row_EW`**, UK: **`r header_row_EW`** 
  - location of the outputs: **`r output_folder`**  

## Plots
  Please check the plots below for any differences between the live data (which 
  is pulled in from the sdg-data (platform) github repo) and the new data.  
  
  Live data will appear over the top of new data. You should therefore only 
  see points for the live data for previous years and only points for the new data 
  in the years you are updating. 

```{r, echo = FALSE, fig.width=10}
live_data <- read.csv('https://raw.githubusercontent.com/ONSdigital/sdg-data/master/data/indicator_3-2-1.csv') %>%
  janitor::clean_names()

new_data <- csv_data %>%
  mutate(Year = as.integer(Year),
         dataset = "new") %>%
  janitor::clean_names() 
  
# join to the new data
joined_data <- live_data %>% 
  mutate(dataset = "live") %>% 
  bind_rows(new_data)

# plot the data
plot <- joined_data %>% 
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  geom_point() +
  geom_line(aes(linetype = dataset)) +
  facet_wrap(vars(gas),
             nrow = 2) +
  theme_bw() 

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

Finally, it is sometimes hard to see where new numbers don't match live numbers
especially if there are lots of disaggregations. In such cases a table of 
differences is also useful. The following example is taken from 13-2-2:
  
  At the time of writing this code (June 2022), the following note is given in the
  source notes tab: 'in these statistics the entire time series going back to 
  1990 is revised each year to take account of methodological improvements'.  
  
  **If there are differences (see table below) please check that this still stands**

```{r, , echo = FALSE}
comparison <- joined_data %>% 
  # select the columns that are in both the live and new data
  # Columns that have the same entry in every row (e.g. units), or
  # columns like geocode and obs status don't need to go in here.
  select(year, gas, value, dataset) %>% 
  pivot_wider(names_from = dataset,
              values_from = value) %>% 
  mutate(difference = new - live) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```
  
  The table below shows all instances where the old value does not match the new 
  value. It is sorted so the largest absolute differences appear first, but can
  be sorted in other ways using the arrows at the top.
```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 5,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

