title: "3-4-2 Checks"
author: "Michael Nairn"
date: "28/07/2023"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(DT)
library(pander)
```



   
## 3-4-2 update summary

This file runs some basic checks and gives information about the update.
  
Before putting the csv in the indicator file please check if the most recent 
figures are provisional. Check this in the source data. If they 
are, in the indicator csv file change the observation status to 'Provisional'.



```{r include= FALSE}
old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/3-4-2.csv") %>%
  mutate(`Drug group` = "Alcohol") %>%
  mutate(Series = recode(Series, "Alcohol - People in treatment" = "People in treatment",
                         "Alcohol - Estimated prevalence" = "Estimated prevalence",
                         "Alcohol - Met need" = "Met need (%)")) %>%
  rename(`Local authority` = Local.Authority) %>%
  mutate(dataset = "old") %>%
  mutate(Sex = "") %>%
  mutate(Age = "") %>%
  mutate(Ethnicity = "") %>%
  select(-c(GeoCode, Unit.multiplier, Unit.measure))

old_data <- old_data %>%
  filter(Deprivation.decile == "") %>%
  filter(!row_number() %in% c(4081)) # random duplication. Won't need in future.
  

date <- Sys.Date()
new_data <- read.csv(paste0("Output/", date, "_3-5-1", ".csv")) %>% 
  rename(`Local authority` = Local.authority) %>%
  rename(`Drug group` = Drug.group) %>%
  mutate(dataset = "new")
 

# join to the new data
joined_data <- old_data %>% 
  bind_rows(new_data) %>%
  select(Year, Series, `Drug group`, `Local authority`, Sex, 
         Age, Ethnicity, Units, Value, dataset)



```



```{r, , echo = FALSE}
comparison <- joined_data %>% 
    select(Year, Series, `Drug group`, `Local authority`, Sex, 
           Age, Ethnicity, Units, Value, dataset) %>% 
    na.omit(Value) %>% # replace the NULL returns with NA
  pivot_wider(names_from = dataset,
              values_from = Value)


comparison <- comparison %>% 
  mutate(across("old":"new", ~replace(., lengths(.) == 0, ""))) %>%
  mutate(new = as.character(new)) %>%
  mutate(new = as.numeric(new)) %>%
  mutate(old = as.character(old)) %>%
  mutate(old = as.numeric(old))

comparison <- comparison %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```

## Table of discrepancies
The table below shows all instances where the old value does not match the new 
value. It is sorted so the largest absolute differences appear first, but can
be sorted in other ways using the arrows at the top.

If there are differences, visible in the table and plots below, please investigate these.
It is possible the data source has update its back time series, so the difference
could be genuine. 


```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```
  



## Plots
Please check the plots below for any differences between the live and new data.
Live data will appear over the top of new data, so if you don't see points for
the new data on dates where there is also live data this is a good thing. 


There are too many possible time series to plot here 
but below are a few examples you can amend if there seems to be a discrepancy. In 
particular for any discrepancies shown in the table. 
You may need to change the filters in the plots. 




```{r, echo = FALSE, fig.width=10}

# People in treatment by drug group
People_in_treatment_headline_plot <- joined_data %>% 
  filter(Series == "People in treatment" & `Local authority` == "" 
         & Ethnicity == "" & Sex == "" & Age == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Series ~ `Drug group`) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("People in treatment by Drug group")


# met need by drug group
met_need_headline_plot <- joined_data %>% 
  filter(Series == "Met need (%)" & `Local authority` == "" 
         & Ethnicity == "" & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Series ~ `Drug group`) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("met need by Drug group")


```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(People_in_treatment_headline_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(met_need_headline_plot))
```


## Session Info
```{r}
pander(sessionInfo())
```

