---
title: "8-8-1 checks"
author: "Katie Uzzell"
date: "05/09/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(pander)
library(DT)
```

```{r include= FALSE}

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/8-8-1.csv")

old_data <- old_data %>% 
  rename("Industry sector" = "Industry.sector", "Units" = "Unit.measure", "Observation status" = "Observation.status", "Unit multiplier" = "Unit.multiplier") 

date <- Sys.Date()

new_data <- read.csv(paste0(output_folder, "/", date, "_update_8-8-1", ".csv"))

latest_year <- max(new_data$Year)

new_data <- new_data %>% 
  rename("Industry sector" = "Industry.sector", "Observation status" = "Observation.status", "Unit multiplier" = "Unit.multiplier") 

new_data <- new_data %>% 
  mutate(dataset = "new") %>% 
  mutate(Value = as.numeric(Value))

# join to the new data
joined_data <- old_data %>% 
  mutate(dataset = "old") %>% 
  bind_rows(new_data) %>% 
  mutate(across(everything(), ~ replace(., is.na(.), ""))) %>% 
  mutate(Value = as.numeric(Value))
```

## 8-8-1 update summary

Data for `r min(new_data$Year)` to `r latest_year` have been created and saved as `r paste0(date, "_8-8-1", ".csv")` in the 8-8-1 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay attention to the time since data were saved, to check that previously downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
8-8-1/`r input_folder`.  

```{r include= FALSE}

fatal_inj_headline_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", fatal_inj_headline), 
                                full.names = TRUE)$ctime[1]
fatal_inj_headline_time_since_data_saved <- round(difftime(Sys.time(), fatal_inj_headline_save_time, units = "days"), 2)

fatal_inj_region_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", fatal_inj_region), 
                                full.names = TRUE)$ctime[1]
fatal_inj_region_time_since_data_saved <- round(difftime(Sys.time(), fatal_inj_region_save_time, units = "days"), 2)

fatal_inj_age_sex_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", fatal_inj_age_sex), 
                                full.names = TRUE)$ctime[1]
fatal_inj_age_sex_time_since_data_saved <- round(difftime(Sys.time(), fatal_inj_age_sex_save_time, units = "days"), 2)

nonfatal_inj_summary_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", nonfatal_inj_summary), 
                                full.names = TRUE)$ctime[1]
nonfatal_inj_summary_time_since_data_saved <- round(difftime(Sys.time(), nonfatal_inj_summary_save_time, units = "days"), 2)

nonfatal_inj_region_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", nonfatal_inj_region), 
                                full.names = TRUE)$ctime[1]
nonfatal_inj_region_time_since_data_saved <- round(difftime(Sys.time(), nonfatal_inj_region_save_time, units = "days"), 2)

nonfatal_inj_age_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", nonfatal_inj_age), 
                                full.names = TRUE)$ctime[1]
nonfatal_inj_age_time_since_data_saved <- round(difftime(Sys.time(), nonfatal_inj_age_save_time, units = "days"), 2)

nonfatal_inj_ind_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", nonfatal_inj_ind), 
                                full.names = TRUE)$ctime[1]
nonfatal_inj_ind_time_since_data_saved <- round(difftime(Sys.time(), nonfatal_inj_ind_save_time, units = "days"), 2)

nonfatal_inj_occ_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", nonfatal_inj_occ), 
                                full.names = TRUE)$ctime[1]
nonfatal_inj_occ_time_since_data_saved <- round(difftime(Sys.time(), nonfatal_inj_occ_save_time, units = "days"), 2)

```


#### **Fatal and non-fatal workplace injuries data from Health and Safety Executive (HSE)**

Fatal workplace injury headline data are saved in the `r fatal_inj_headline`. This file was saved at `r fatal_inj_headline_save_time`: It has been 
`r fatal_inj_headline_time_since_data_saved` days since the file was saved in `r input_folder`. 
 
Fatal workplace injury by region data are saved in the `r fatal_inj_region`.This file was saved at `r fatal_inj_region_save_time`: It has been 
`r fatal_inj_region_time_since_data_saved` days since the file was saved in `r input_folder`. 

Fatal workplace injury by age and sex data are saved in the `r fatal_inj_age_sex`. This file was saved at `r fatal_inj_age_sex_save_time`: It has been 
`r fatal_inj_age_sex_time_since_data_saved` days since the file was saved in `r input_folder`. 

Non-fatal workplace injury headline data data are saved in the `r nonfatal_inj_summary`. This file was saved at `r nonfatal_inj_summary_save_time`: It has been 
`r nonfatal_inj_summary_time_since_data_saved` days since the file was saved in `r input_folder`. 

Non-fatal workplace injury by region data are saved in the `r nonfatal_inj_region`. This file was saved at `r nonfatal_inj_region_save_time`: It has been 
`r nonfatal_inj_region_time_since_data_saved` days since the file was saved in `r input_folder`. 

Non-fatal workplace injury by age and sex data are saved in the `r nonfatal_inj_age`. This file was saved at `r nonfatal_inj_age_save_time`: It has been 
`r nonfatal_inj_age_time_since_data_saved` days since the file was saved in `r input_folder`. 

Non-fatal workplace injury by industry data are saved in the `r nonfatal_inj_ind`. This file was saved at `r nonfatal_inj_ind_save_time`: It has been 
`r nonfatal_inj_ind_time_since_data_saved` days since the file was saved in `r input_folder`. 

Non-fatal workplace injury by occupation data are saved in the `r nonfatal_inj_occ`. This file was saved at `r nonfatal_inj_occ_save_time`: It has been 
`r nonfatal_inj_occ_time_since_data_saved` days since the file was saved in `r input_folder`. 


**The tab used for all fatal injury inputs is called `r fatal_tabname`.**

**The tab used for all non-fatal injury inputs is called `r nonfatal_tabname`.**

## **Check data**

Please check the table below for any differences between the live and new data. If there are differences visible in the table below, please investigate these.

```{r include=FALSE}

comparison <- joined_data %>% 
  select(Year, Series, Age, Sex, Country, Region, `Industry sector`, Occupation, Value, dataset) %>%
  pivot_wider(names_from = dataset,
              values_from = Value)

comparison <- comparison %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```


```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

### Calculations

No calculations were performed on this data during the automated update. 

## Session Info
```{r}
pander(sessionInfo())
```

