title: "3-b-1 Checks"
author: "Michael Nairn"
date: "15/08/2023"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(DT)
library(pander)
```



   
## 3-b-1 update summary

This file runs some basic checks and gives information about the update.

This automation only updates Source 1. Source 2 is discontinued so is not updated by this automation. Make sure not to overwrite source 2 data when putting the automated csv file into csv sheet of indicator file.

Before putting the csv in the indicator file please check if the most recent 
figures are provisional. Check this in the source data. If they 
are, in the indicator csv file change the observation status to 'Provisional'.



```{r include= FALSE}
old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/3-b-1.csv") %>%
  rename(Units = Unit.measure) %>%
    mutate(dataset = "old") %>%
    select(Year, Country, Region, Sex, Age, Units, Value, dataset)


date <- Sys.Date()
new_data <- read.csv(paste0("Output/", date, "_3-b-1", ".csv")) %>% 
  mutate(dataset = "new") %>%
  mutate(Value = as.numeric(Value)) %>%
  select(Year, Country, Region, Sex, Age, Units, Value, dataset)
 

# join to the new data
joined_data <- old_data %>% 
  bind_rows(new_data)

```



```{r, , echo = FALSE}
comparison <- joined_data %>%
  na.omit(Value) %>% # replace the NULL returns with NA
  pivot_wider(names_from = dataset,
              values_from = Value)


comparison <- comparison %>% 
  mutate(across("old":"new", ~replace(., lengths(.) == 0, ""))) %>%
  mutate(new = as.character(new)) %>%
  mutate(new = as.numeric(new)) %>%
  mutate(old = as.character(old)) %>%
  mutate(old = as.numeric(old))

comparison <- comparison %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```

## Table of discrepancies
The table below shows all instances where the old value does not match the new 
value. It is sorted so the largest absolute differences appear first, but can
be sorted in other ways using the arrows at the top.

If there are differences, visible in the table and plots below, please investigate these.
It is possible the data source has update its back time series, so the difference
could be genuine. Please see source notes, suicide details can be reassigned
as polics investigations proceed.


```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```
  



## Plots
Please check the plots below for any differences between the live and new data.
Live data will appear over the top of new data, so if you don't see points for
the new data on dates where there is also live data this is a good thing. 


There are too many possible time series to plot here 
but below are a few examples you can amend if there seems to be a discrepancy. In 
particular for any discrepancies shown in the table. 
You may need to change the filters in the plots. 



```{r, echo = FALSE, fig.width=10}

# Sex headline plot
sex_headline_plot <- joined_data %>% 
  filter(Region == "" & Age == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Country ~ Sex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Suicides by Country and sex")

# Age headline plot
age_headline_plot <- joined_data %>% 
  filter(Region == "" & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Country ~ Age) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Suicides by Country and age")



```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(sex_headline_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(age_headline_plot))
```

```{r, echo = FALSE, fig.width=10}

# Sex headline plot
sex_region_plot <- joined_data %>% 
  filter(Region != "" & Age == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Region ~ Sex) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Suicides by Region and sex")

# Age headline plot
age_region_plot <- joined_data %>% 
  filter(Region != "" & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(Region ~ Age) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7)) +
  ggtitle("Suicides by region and age")



```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(sex_region_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(age_region_plot))
```


## Session Info
```{r}
pander(sessionInfo())
```

