---
title: "8-10-2 checks"
author: "Katie Uzzell"
date: "25/04/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
```

```{r include= FALSE}

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/8-10-2.csv")

date <- Sys.Date()
new_data <- read.csv(paste0(output_folder, "/", date, "_update_8-10-2", ".csv"))

latest_year <- max(new_data$Year)

```

## 8-10-2 update summary

Data for `r latest_year` have been created and saved as `r paste0(date, "_8-10-2", ".csv")` 
in the 8-10-2 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay 
attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
8-10-2/`r input_folder`.  
  
```{r include= FALSE}
filename_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", filename), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved <- round(difftime(Sys.time(), filename_save_time, units = "days"), 2)

```

**Savings and investment data from the Family Resources Survey (DWP)**  

Savings and investment data are in the file `r filename`.   

This file was saved at `r filename_save_time`: It has been 
`r time_since_data_saved` days since the file was saved in 
`r input_folder`.  

The tab used is called
`r tabname`. 

### Check data

Because this automation only produces the data for the latest year, it is not 
possible to plot the new data over the old data to check for discrepancies. 
Instead, the person QAing this update should download the source data and spot 
check values the source data against the outputted csv.

### Calculations

No calculations were performed on this data.

