title: "3-4-1 QA"
author: "Michael Nairn"
date: "21/11/2022"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(DT)
library(pander)
```



   
## 3-4-1 update summary

3-4-1 is an indicator that uses a NOMIS API link to provide data. Unless the 
format of NOMIS API downloads changes, the update_3-4-1 script should not need 
changing. However, always check before running the update script. 

This file runs some basic checks and gives information about the update.
  
Before putting the csv in the indicator file please check if the most recent 
figures are provisional. Check this in the source data. If they 
are, in the indicator csv file change the observation status to 'Provisional'.


Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_3-4-1", ".csv")` in the 3-4-1 Output folder.  



```{r include= FALSE}
source('config.R')

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/3-4-1.csv") %>%
  rename("Type of disease" = "Type.of.disease",
         "Chronic respiratory disease subtype" = "Chronic.respiratory.disease.subtype")

date <- Sys.Date()
new_data <- read.csv(paste0("3-4-1_output/", date, "_3-4-1", ".csv")) %>% 
  mutate(Year = as.integer(Year),
         dataset = "new") %>%
  rename("Type of disease" = "Type.of.disease",
         "Chronic respiratory disease subtype" = "Chronic.respiratory.disease.subtype") 

latest_year <- max(new_data$Year)

# join to the new data
joined_data <- old_data %>% 
  mutate(dataset = "old") %>% 
  bind_rows(new_data) 
```


## Plots
Please check the plots below for any differences between the live and new data.
Live data will appear over the top of new data, so if you don't see points for
the new data on dates where there is also live data this is a good thing.  


```{r, echo = FALSE, fig.width=10}

# Country by type of disease
country_headline_plot <- joined_data %>% 
  filter(Region == "" & `Chronic respiratory disease subtype` == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(`Type of disease` ~ Country) +
  theme_bw() +
  ggtitle("Country by Type of disease")


# Age by type of disease
age_headline_plot <- joined_data %>% 
  filter(`Chronic respiratory disease subtype` == "" & Region == "" & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_grid(`Type of disease` ~ Age) +
  theme_bw() +
  ggtitle("Age by Type of disease")


```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(country_headline_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(age_headline_plot))
```

  
**If there are differences (see table below) please check that this still stands**

{r, , echo = FALSE}
comparison <- joined_data %>% 
  select(Year, Gas, Value, dataset) %>% 
  pivot_wider(names_from = dataset,
              values_from = Value) %>% 
  mutate(difference = new - live) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))


  
The table below shows all instances where the old value does not match the new 
value. It is sorted so the largest absolute differences appear first, but can
be sorted in other ways using the arrows at the top.
{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))


## Session Info
```{r}
pander(sessionInfo())
```

