title: "3-4-1 QA"
author: "Michael Nairn"
date: "21/11/2022"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(DT)
library(pander)
```



   
## 3-4-1 update summary

3-4-1 is an indicator that uses a NOMIS API link to provide data. Unless the 
format of NOMIS API downloads changes, the update_3-4-1 script should not need 
changing. However, always check before running the update script. 

This file runs some basic checks and gives information about the update.
  
Before putting the csv in the indicator file please check if the most recent 
figures are provisional. Check this in the source data. If they 
are, in the indicator csv file change the observation status to 'Provisional'.



```{r include= FALSE}
old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/3-4-1.csv") %>%
  rename("Type of disease" = "Type.of.disease",
         "Chronic respiratory disease subtype" = "Chronic.respiratory.disease.subtype") %>%
  mutate(dataset = "old") %>%
  select(-GeoCode)
  

date <- Sys.Date()
new_data <- read.csv(paste0("3-4-1_output/", date, "_3-4-1", ".csv")) %>% 
  mutate(Year = as.integer(Year),
         dataset = "new") %>%
  rename("Type of disease" = "Type.of.disease",
         "Chronic respiratory disease subtype" = "Chronic.respiratory.disease.subtype") 

latest_year <- max(new_data$Year)

# join to the new data
joined_data <- old_data %>% 
  bind_rows(new_data) %>%
  rename(`Observation status` = `Observation.status`,
         `Unit multiplier` = `Unit.multiplier`,
         `Unit measure` = `Unit.measure`)

```


## Plots
Please check the plots below for any differences between the live and new data.
Live data will appear over the top of new data, so if you don't see points for
the new data on dates where there is also live data this is a good thing.  


```{r, echo = FALSE, fig.width=10}

# Country by type of disease
country_headline_plot <- joined_data %>% 
  filter(Region == "" & `Chronic respiratory disease subtype` == "" & Age == ""
         & Sex == "" ) %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Type of disease` ~ Country) +
  theme_bw() +
  ggtitle("Country by Type of disease")


# Region by type of disease
region_headline_plot <- joined_data %>% 
  filter(Country =="England" & `Chronic respiratory disease subtype` == "" & Age == ""
         & Sex == "" ) %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Type of disease` ~ Region) +
  theme_bw() +
  ggtitle("Region by Type of disease")


# Age by type of disease
age_headline_plot <- joined_data %>% 
  filter(`Chronic respiratory disease subtype` == "" & Region == "" & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Type of disease` ~ Age) +
  theme_bw() +
  ggtitle("Age by Type of disease")


# Sex by type of disease
sex_headline_plot <- joined_data %>% 
  filter(Region == "" & `Chronic respiratory disease subtype` == "" & Age == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Type of disease` ~ Sex) +
  theme_bw() +
  ggtitle("Sex by Type of disease")

```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(country_headline_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(region_headline_plot))
```


```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(age_headline_plot))
```

  
```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(sex_headline_plot))
```


```{r, echo = FALSE, fig.width=10}

# Country by chronic respiratory disease subtype
chronic_respiratory_disease_plot <- joined_data %>% 
  filter(Region == "" & `Type of disease` == "Chronic respiratory disease" & Age == ""
         & Sex == "" ) %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Chronic respiratory disease subtype` ~ Country) +
  theme_bw() +
  ggtitle("Country by chronic respiratory disease subtype")

# Region by chronic respiratory disease subtype
region_chronic_respiratory_disease_plot <- joined_data %>% 
  filter(Country == "England" & `Type of disease` == "Chronic respiratory disease" & Age == ""
         & Sex == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Chronic respiratory disease subtype` ~ Region) +
  theme_bw() +
  ggtitle("Region by chronic respiratory disease subtype")


# Age by chronic respiratory disease subtype
age_chronic_respiratory_disease_plot <- joined_data %>% 
  filter(`Type of disease` == "Chronic respiratory disease" & Region == ""
         & Sex == "" ) %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Chronic respiratory disease subtype` ~ Age) +
  theme_bw() +
  ggtitle("Age by chronic respiratory disease subtype")


# Sex by chronic respiratory disease subtype
sex_chronic_respiratory_disease_plot <- joined_data %>% 
  filter(`Type of disease` == "Chronic respiratory disease" & Region == ""
         & Age == "") %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  facet_grid(`Chronic respiratory disease subtype` ~ Sex) +
  theme_bw() +
  ggtitle("Sex by chronic respiratory disease subtype")

```


```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(chronic_respiratory_disease_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(region_chronic_respiratory_disease_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(age_chronic_respiratory_disease_plot))
```

```{r, echo = FALSE, fig.width=10, fig.height = 10}
suppressMessages(print(sex_chronic_respiratory_disease_plot))
```



If there are differences, visible in the table below and plots, please investigate these.
It is possible the data source has update its back time series, so the difference
could be genuine. 

{r, , echo = FALSE}
comparison <- joined_data %>% 
  select(Year, Country, Region, Age, Sex, `Type of disease`,
         `Chronic respiratory disease subtype`, Value, dataset) %>%
    na.omit(Value)

comparison <- pivot_wider(comparison, names_from = dataset,
              values_from = Value) %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))



The table below shows all instances where the old value does not match the new 
value. It is sorted so the largest absolute differences appear first, but can
be sorted in other ways using the arrows at the top.

```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```
  


## Session Info
```{r}
pander(sessionInfo())
```

