---
title: "QA 3-9-1"
author: "Michael Nairn"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
library(pander)
```
## Introduction

3-9-1 is one of the indicators on Fingertips. The source can be a pain to locate,
and it is very straightforward to update. It used to be automated using Python,
but that code is now in the archive as it used an out of date link.  
  
The method for this PHE Public Health Outcome Framework indicator has recently
been changed but they do not intend to use the new method on the backseries. The 
two series should not be presented on the same chart without making this clear.
   
This file runs some basic checks on the csv output and highlights some specific 
things that should manually be checked.
  
Before putting the csv in the indicator file please check in the source data if the most recent 
figures are provisional. If they are, in the indicator csv file change the 
observation status to 'Provisional'.
  
 
## Configurations

  The configurations are:  
  
  - filepath for the input data: **`r paste0(getwd(), "/", input_folder, "/", csv_filename)`**
  
  

## Plots
Please check the plots below for any differences between the live data (which 
is pulled in from the live site) and the new data. 

Local authority data is not plotted due to there being too many to visualise. 
Please use the table of differences to guide investigation of differences in local authority data.
  
Live data will appear over the top of new data. You should therefore only 
see points for the live data for previous years and only points for the new data 
in the years you are updating. 

```{r, echo = FALSE}
# functions for making tables & plots of differences between live and new data
# plots:
filter_for_plot <- function(series_value, var1, var2) {
  
  standard_cols <- c("year", "value", "units", "series",
                     "observation_status", "unit_multiplier",
                     "dataset", var1, var2)
  
  series_filter <- joined_data %>% 
    filter(series == series_value)
  
  # get just the columns we want to filter to be blank
  disaggs <- setdiff(names(joined_data), standard_cols)
  
  for (i in 1:length(disaggs)) {
    series_filter <- series_filter %>% 
      filter(!!(as.name(disaggs[i])) == "" | is.na(!!(as.name(disaggs[i]))))
  }
  
  return(series_filter)
} 
make_plot <- function(dat, var1, var2) {
  
  ggplot(dat,
         aes(x = year,
             y = value,
             colour = dataset)) +
    geom_point() +
    geom_line() +
    facet_grid(vars({{var1}}), vars({{var2}})) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45,
                                     vjust = 1, hjust = 1))
}
# tables
calculate_differences <- function(dat, var1, var2) {
  
  dat %>% 
    mutate(disagg_combo = paste(!!as.name(var1), ",", !!as.name(var2))) %>% 
    select(year, series, disagg_combo,
           dataset, value)  %>% 
    pivot_wider(names_from = dataset,
                values_from = value) %>% 
    mutate(difference = new - live) %>% 
    filter(difference != 0) %>%
    # put the largest differences at the top
    arrange(desc(abs(difference)))
  
}
# and then to present it as a datatable
create_datatable <- function (data) {
  
  datatable(data, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 3,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
}
```

```{r, echo = FALSE, fig.width=10}

# do this properly once new data are live - at the moment the two are too
# different to code easily. Figures that are calculated the same do match.

live_data <- read.csv('https://raw.githubusercontent.com/ONSdigital/sdg-data/master/data/indicator_3-9-1.csv') %>%
  janitor::clean_names() %>% 
  mutate(dataset = "live") %>% 
  mutate(across(where(is.character), str_squish)) 

# local authority is different in the new data

new_data <- read.csv(paste0(output_folder, "/", csv_filename)) %>%
  janitor::clean_names() %>% 
  mutate(dataset = "new") %>% 
  mutate(across(where(is.character), str_squish)) 

# join to the new data
joined_data <- live_data %>% 
  bind_rows(new_data) 

# plot the data
joined_data %>% 
  filter(#(is.na(region) | region == "") & 
           (is.na(local_authority) | local_authority == "") & 
           (is.na(deprivation_decile) | deprivation_decile == "")) %>%  
  ggplot(.,
         aes(x = year,
             y = value,
             colour = dataset)) +
  scale_x_continuous (breaks=min (joined_data$year):max (joined_data$year)) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(region),
             nrow = 2) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7))

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

  
  **If there are differences (see table below) please check that this still stands**

```{r, echo = FALSE}
comparison <- joined_data %>% 
  pivot_wider(names_from = dataset,
              values_from = value)

comparison <- comparison %>% 
  mutate(difference = new - live) %>%
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```


  The table below shows all instances where the old value does not match the new 
  value. It is sorted so the largest absolute differences appear first, but can
  be sorted in other ways using the arrows at the top.
  
```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

pander(sessionInfo())
