---
title: "QA 3-9-1"
author: "Emma Wood"
date: `r Sys.time`
output: html_document
---
*This is only a template and will not produce real output as the variables *
*referred to in the code do not exist - in a real one they will be pulled from*
*the other scripts. For a working QA file similar to this template see 13-2-2*

*Example text is indented, but shouldn't be indented in the real thing*

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(DT)
```
## Introduction

3-9-1 is one of the indicators on Fingertips. The source can be a pain to locate,
and it is very straightforward to update. It used to be automated using Python,
but that code is now in the archive as it used an out of date link.  
  
The method for this PHE Public Health Outcome Framework indicator has recently
been changed but they do not intend to use the new method on the backseries. The 
two series should not be presented on the same chart without making this clear
   
This file runs some basic checks on the csv output and highlights some specific 
things that should manually be checked.
  
- specific things that should be checked in addition to reading this document 
e.g. if provisional figures are likely to be included but this info is not 
picked up by the automation you may put something like "Before putting the csv 
in the indicator file please check in the source data if the most recent 
figures are provisional. If they are, in the indicator csv file change the 
observation status to 'Provisional'."  
  
- other important info you think the user needs to be aware of.
  
## Configurations
in this section list the settings (configurations) and use snippets of code to 
print the user input. The double asterisks make the text bold, while `r` tells 
markdown to read the text inside the back-ticks as code. 

An example of what you may put in this section -
  
  The configurations are:  
  
  - filepath for the input data: **`r paste0(getwd(), "/", input_folder, "/", filename)`**
  - name of the tab the data are in: **`r tab_name`**  
  - row number of the column names: **`r header_row`** 
  - location of the outputs: **`r output_folder`**  
  
  
## Basic checks
Use this section to highlight anything you think may have gone wrong in the 
code. You may want to use chunks of code within this file to perform some basic 
checks, or you may just pull in variables that were created in the main scripts 
(like with the configurations above). 

For example, when pulling data from nomis, where there are multiple year
types (e.g. Jan to Dec, Apr to March etc), it may not be clear just from the 
output which data have been pulled from nomis. In such a case this section may read:

  We should only use data for the calendar year (Jan-Dec). The months contained in 
  the data for this run are: `r unique_months_retained`.  If the given months are 
  incorrect please check the nomis selections are correct, and regenerate the link 
  (see the 3-c-1 README file).  

You can create a statement that is printed in the file only if a certain 
condition has been met. For example, in 2-c-1 two datasets feed into the csv, 
and they are released at different times. You may want to therefor make the 
person doing the update aware that one dataset is not yet released:

```{r, include=FALSE}

# # in this example from 2-c-1, not_all_data_available is a variable created in 
# # the main code (copied and hashed out below for demonstration only).
# # not_all_data_available is TRUE if the most recent year of one dataset was 
# # different to the most recent year in the other dataset, and FALSE if they 
# # both go up to the same year:
#
# if (length(unique(max_years$max_year)) > 1) {
#   not_all_data_available <- TRUE
#   dataset_behind <- max_years$dataset[max_years$max_year == full_data_max_year]
# } else {
#   not_all_data_available <- FALSE
# }

if (not_all_data_available == TRUE) {
  print(paste("NOTE: Data for", dataset_behind, "data are only available up to", 
              max_years$max_year[max_years$dataset == dataset_behind], 
              ", though data for", max_years$dataset[max_years$dataset != dataset_behind],
              "are available up to", max_years$max_year[max_years$dataset != dataset_behind]))
}
```
  
In other updates, this section may be much more basic, for example just checking 
that the geography given in the table title matches the geography in the output. 
For example you may include something like:

  The geography listed in the information at the top of 
  `r paste0(filename, ", table ", tab_name)` is `r country`. Please check that 
  this matches the geography given in the indicator file metadata.  
     
Sometimes, there may not be anything that needs to go in here, in which case 
feel free to leave the section out.  

## Plots
The plots are one of the most important parts of the QA file. If anything has 
gone wrong in the automation it will show up on the plots. Below is an example
plots section:  
  
  Please check the plots below for any differences between the live data (which 
  is pulled in from Jemalex) and the new data.  
  
  Live data will appear over the top of new data. You should therefore only 
  see points for the live data for previous years and only points for the new data 
  in the years you are updating. 

```{r, echo = FALSE, fig.width=10}

# This example is taken from the 13-2-2 update because it is a straightforward
# one.
#
# Some will be more complex, for example when there are multiple disaggregations.
# See the 3-c-1 QA.Rmd file for an example of more complex QA plots
#
# Note that you may have improved the indicator at the same time as updating it
# when writing the automation. If this is the case, the plots code you write for 
# the first round of checks may need to be slightly different to how it will 
# need to be in future years. So you will need to check the Rmd file works
# with 'live' data that looks like the new data (for testing you could pull the 
# 'live' data from the in progress file once you have put the new csv in there).

live_data <- read.csv('https://raw.githubusercontent.com/ONSdigital/sdg-data/master/data/indicator_13-2-2.csv') %>%
  janitor::clean_names()

new_data <- csv_formatted %>% 
  mutate(Year = as.integer(Year),
         dataset = "new") %>%
  janitor::clean_names() 
  
# join to the new data
joined_data <- live_data %>% 
  mutate(dataset = "live") %>% 
  bind_rows(new_data)

# plot the data
plot <- joined_data %>% 
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(Gas),
             nrow = 2) +
  theme_bw() 

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

Finally, it is sometimes hard to see where new numbers don't match live numbers
especially if there are lots of disaggregations. In such cases a table of 
differences is also useful. The following example is taken from 13-2-2:
  
  At the time of writing this code (June 2022), the following note is given in the
  source notes tab: 'in these statistics the entire time series going back to 
  1990 is revised each year to take account of methodological improvements'.  
  
  **If there are differences (see table below) please check that this still stands**

```{r, , echo = FALSE}
comparison <- joined_data %>% 
  select(Year, Gas, Value, dataset) %>% 
  pivot_wider(names_from = dataset,
              values_from = Value) %>% 
  mutate(difference = new - live) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```
  
  The table below shows all instances where the old value does not match the new 
  value. It is sorted so the largest absolute differences appear first, but can
  be sorted in other ways using the arrows at the top.
```{r, echo = FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

pander(sessionInfo())
