---
title: "3-a-1 checks"
author: "Katie Uzzell"
date: "22/02/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r echo=FALSE}
load('img.RData')
```

```{r load-packages, echo=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(DT)
library(pander)
```

```{r echo=FALSE}
old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/3-a-1.csv") %>%
  mutate(dataset = "old") %>%
  select(-GeoCode)

date <- Sys.Date()
new_data <- read.csv(paste0(output_folder, "/", date, "_3-a-1", ".csv"))

latest_year <- max(new_data$Year)
```
## **3-a-1 update summary**

The 3-a-1 automation uses multiple tabs of one excel spreadsheet to report data
on 'Percentage of people who are current cigarette smokers aged 18 years and older'
disaggregated by sex, age, country, England region, country of birth, socio-economic
classification, and ethnicity.

This file runs some basic checks and gives information about the update.

Before putting the csv in the indicator file please check if the most recent 
figures are provisional. Check this in the source data. If they 
are, in the indicator csv file change the observation status to 'Provisional'.
Please also check the source data for any metadata that may also need to be included
on the platform.

Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_3-a-1", ".csv")` in the 3-a-1 Output folder.  

### **Check configurations**

Please check that the information below is as you expect. In particular, pay 
attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.

These settings can all be adjusted in `config.R`

Files used to create the new data were downloaded and saved in 
3-a-1/`r input_folder`.  

```{r echo=FALSE}
filename_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", filename), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved <- round(difftime(Sys.time(), filename_save_time, units = "days"), 2)
```

```{r include=FALSE}
new_data <- new_data %>%
  mutate(dataset = "new")

# join to the new data
joined_data <- old_data %>% 
  bind_rows(new_data) 
```

### **Smoking habit data from health data team (ONS)**  

Smoking habit data are in the file `r filename`.   

This file was saved at `r filename_save_time`: It has been 
`r time_since_data_saved` days since the file was saved in 
`r input_folder`  

The tabs used are called
`r tabnames`. 

### **Calculations**

No calculations were performed on this data.

## **Check data**

Please check the table below for any differences between the live and new data.
If there are differences visible in the table below, please investigate these.
It is possible the data source has update its back time series, so the difference
could be genuine. 

```{r include=FALSE}

comparison <- joined_data %>% 
  select(Year, Sex, Age, Country, Region, Socio.economic.classification, Ethnicity, Country.of.birth,
         Value, dataset) %>%
  pivot_wider(names_from = dataset,
              values_from = Value)

comparison <- comparison %>% 
  mutate(difference = new - old) %>% 
  filter(difference != 0) %>% 
  # put the largest differences at the top
  arrange(desc(abs(difference)))

```

The table below shows all instances where the old value does not match the new 
value. It is sorted so the largest absolute differences appear first, but can
be sorted in other ways using the arrows at the top.

```{r echo=FALSE}
datatable(comparison, 
          rownames = FALSE, 
          filter = "top",
          options = list(dom = 'tp', 
                         pageLength = 10,
                         columnDefs = list(list(
                           className = 'dt-center', targets = 0:3))))
```

## Session Info
```{r}
pander(sessionInfo())
```

