---
title: "8-1-1 checks"
author: "Katie Uzzell"
date: "03/05/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}
library(tidyr)
library(dplyr)
library(ggplot2)
library(kableExtra)
library(pander)
```

```{r include= FALSE}

old_data <- read.csv("https://sdgdata.gov.uk/sdg-data/en/data/8-1-1.csv")

date <- Sys.Date()
new_data <- read.csv(paste0(output_folder, "/", date, "_update_8-1-1", ".csv"))

latest_year <- max(new_data$Year)

```

## 8-1-1 update summary

Data for `r min(new_data$Year)` to `r latest_year` have been created and 
saved as `r paste0(date, "_8-1-1", ".csv")` in the 8-1-1 Output folder.  
    
### Check configurations
Please check that the information below is as you expect. In particular, pay 
attention to the time since data were saved, to check that previously 
downloaded data have not accidentally been used.
    
These settings can all be adjusted in `config.R`
    
Files used to create the new data were downloaded and saved in 
8-1-1/`r input_folder`.  
  
```{r include= FALSE}
filename_save_time <- file.info(paste0(getwd(),"/",input_folder,"/", filename), 
                                full.names = TRUE)$ctime[1]
time_since_data_saved <- round(difftime(Sys.time(), filename_save_time, units = "days"), 2)

```

**GDP data from UK environmental accounts time series dataset (ONS)**  

GDP data are in the file `r filename`. The time series used for this indicator is N3Y6 (GDP (Average) per head, year on year growth rate CVM SA %) located in column DLH of the source download.  

This file was saved at `r filename_save_time`: It has been 
`r time_since_data_saved` days since the file was saved in 
`r input_folder`  

The tab used is called
`r tabname`. 


### Check data
The charts below show the new data (just created) plotted on top of the old data, which has been downloaded from the live site. It is likely there will have been revisions to the backseries, so small discrepancies are expected. If there are unexpected or large discrepancies between the old and new data, please investigate the reasons for these discrepancies.
    
```{r, echo=FALSE, fig.width=7, fig.height=15}
new_data <- new_data %>%
mutate(Year = as.integer(Year),
         dataset = "new")

new_data$Value <- as.numeric(new_data$Value)

joined_data <- old_data %>%
  mutate(dataset = "old") %>%
  bind_rows(new_data)
 
plot <- joined_data %>%
  ggplot(.,
         aes(x = Year,
             y = Value,
             colour = dataset)) +
    #scale_x_continuous(breaks = seq(min(year(joined_data$Year)), max(year(joined_data$Year)), by = 3))+
  scale_x_continuous (breaks=min (joined_data$Year):max (joined_data$Year)) +
  geom_point() +
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 7)) +
  theme(axis.text.y = element_text(size = 7))

```


```{r, echo = FALSE, fig.height = 8, fig.width = 10}
suppressMessages(print(plot))
```

### Calculations

No calculations were performed on this data.

## Session Info
```{r}
pander(sessionInfo())
```

